### 2016-12-07 Reports/2016_12_05/Reg_age_severity_simulation.Rmd
make plots
make plots

---
title: "Reg_age_severity_simulation"
author: "Detian Deng"
date: "December 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r warning=FALSE, message=FALSE}
setwd("~/Documents/JHSPH/Research/S.Zeger/PQ_Models/Code")
source("./PerchDataSimulation.R")
source("./ModelFittingJAGS.R")
source("./PostProcessJAGS.R")
source("./PerchPlots.R")

library(corrplot)
load("~/Documents/JHSPH/Research/S.Zeger/PQ_Models/WorkSpace/SC1-2CovWithInteract-theta2--4-tau-0.2.RData")
```

## Descriptive Analysis
```{r }
X.unique = uniquecombs(sim.obj$X)
strata.id = attr(X.unique, "index")
X.unique = matrix(X.unique, nrow(X.unique))
rownames(X.unique) = paste("strata", 1:nrow(X.unique))
colnames(X.unique) = c("intercept", "age", "severity", "age*severity")
X.unique
table(strata.id)
```

### SS data: Count by Strata
```{r}
mss.tab = aggregate(sim.obj$MSS.case, list(strata.id), sum)
colnames(mss.tab) = c("strata", "Pathogen_A", "Pathogen_B", "Pathogen_C",
                      "Pathogen_D", "Pathogen_E")
mss.tab
```

### BS data: Count by Strata
```{r}
mbs.tab = aggregate(sim.obj$MBS.case, list(strata.id), sum)
colnames(mbs.tab) = c("strata", "Pathogen_A", "Pathogen_B", "Pathogen_C",
                      "Pathogen_D", "Pathogen_E")
mbs.tab
```

## Marginal Etiology by Strata
```{r }
PlotByPathogen(coda.fit[[1]], sim.obj)
```

## Etiology Cell Probabilities - Posterior vs. Prior
```{r}
PlotByCombination(coda.fit[[1]], sim.obj, hyper.pars.list)
```

## Chain Correlaiton
```{r fig.width=25, fig.height=20}
par.all = SweepPosterior(coda.fit[[1]], sim.obj, sweep.on = NULL,
                         model.type = "regression")

# par.swept = SweepPosterior(coda.fit[[1]], sim.obj, sweep.on = "ss_tpr",
#                            model.type = "regression")

C = cor(par.all)
corrplot.mixed(C, lower = "number", upper = "ellipse")
```